Welcome to the \lmc\ User's Guide!

In this User's Guide we describe how to download and run \lmc, a
massively parallel code that solves the multicomponent hydrodynamic equations for
chemically reacting flows in the low Mach number regime.  \lmc\ uses an Eulerian grid and
incorporates adaptive mesh refinement (AMR).  Our approach to AMR uses
a nested hierarchy of logically-rectangular grids with simultaneous
refinement in both space and time, utilizing the \boxlib\ library. 

The core algorithms in \lmc\ are described in a series of papers:
\begin{itemize}
\item {\it Numerical Simulation of Laminar Reacting Flows with Complex Chemistry}, M.~S.~Day and J.~B.~Bell,
  Combust. Theory Modelling 4(4) pp.535-556, 2000.
  \url{https://ccse.lbl.gov/Publications/day/db99/db99.pdf}

\item {\it An Adaptive Projection Method for Unsteady, Low-Mach Number Combustion}, 
  R.~B.~Pember, L.~H.~Howell, J.~B.~Bell, P.~Colella, W.~Y.~Crutchfield, W.~A.~Fiveland, and J.~P.~Jessee,
  Comb. Sci. Tech., 140, pp. 123-168, 1998.

\item {\it A Conservative Adaptive Projection Method for the Variable Density Incompressible Navier-Stokes Equations},
  A.~S.~Almgren, J.~B.~Bell, P.~Colella, L.~H.~Howell, and M.~L.~Welcome, J.~Comp.~Phys., 142, pp. 1-46, 1998.
  \url{https://ccse.lbl.gov/Publications/almgren/abchw96/paper.ps.gz}

\end{itemize}         

The development of \lmc\ and the \boxlib\ library is led by the Center for Computational
Sciences and Engineering / Lawrence Berkeley National Laboratory (\url{ccse.lbl.gov})

Past and present contributors to the \lmc\ code base include: % git shortlog -s -n
\begin{quote}
Ann Almgren\newline
Andy Aspden\newline
John Bell\newline
Vince Beckner\newline
Marc Day\newline
Matthew Emmett\newline
Candace Gilet\newline
Ray Grout\newline
Mike Lijewski\newline
Andy Nonaka\newline
Will Pazner\newline
Chuck Rendleman\newline
Weiqun Zhang\newline
\end{quote}

TODO: git repo, contacts, mailing list
